---
title       : DATA MINING-FINAL PROJECT 
subtitle    : DIABETES DETECTION
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---


## PROJECT OVERVIEW

1. Data Mining project to predict Diabetes in Females participants
2. Binomial Regression 
3. Comes under classification with yes/no output

---

## DATA INFORMATION

1. Data contains 768 instances all female participants
2. Has 8 predictor attributes and 1 predicted attribute 
3. Explaining attributes

-Pregnancies-Number of pregnancies(Numeric)

-Glucose-Plasma glucose concentration (Numeric)
    
-BloodPressure-Diastolic blood pressure (mm Hg (Numeric)	
    
-SkinThickness-Triceps skin fold thickness (mm) (Numeric)	
    
-Insulin- 2-Hour serum insulin (mu U/ml) (Numeric)	
    
-BMI- Body mass index level (Numeric)	
    
-DPF-	Diabetic Pedgree Factor (Numeric) Ancestorial history of disease

-Age- Age (years) (numeric)

-OutcomeClass variable (YES or NO)

---

## libraries required
```{r}
library(caret)
library(randomForest)
library(e1071)
library(klaR)
library(e1071)
library(corrplot)
library(ggplot2)
library(hexbin)
library(slidify)
library(slidifyLibraries)
library(lattice)
```

---

## Loading the data for exploratory data analysis
```{r}
diabetes<- read.csv("C:/Users/Madan/Documents/diabetes.csv",header=TRUE, stringsAsFactors = TRUE)
diabetes
```

---

## EDA-PLOT-Detected and non-detected
```{r}
library(ggplot2)
plot1=ggplot(diabetes, aes(x =diabetes$Outcome )) + geom_bar()

```

---

```{r}
plot1
```

---

## EDA-CORRELATION PLOT-All variables
```{r}
library(corrplot)
diabetes1<- read.csv("C:/Users/Madan/Documents/diabetes1.csv",header=TRUE, stringsAsFactors = TRUE)
corplot=cor(diabetes1)
```

---

```{r}
corplot1=corrplot(corplot,order="hclust",method="number")
```

---

## EDA-SCATTER-Age vs Pregnancies
```{r}
scatter1=xyplot(diabetes$Age ~ diabetes$Pregnancies, groups=diabetes$Outcome, data = diabetes,auto.key = list(corner = c(0, .98)), cex = .75)
```

---

```{r}
scatter1 ##age vs preg
```

---

```{r}
d <- ggplot(diabetes, aes(Pregnancies,Age))
```

---

```{r}
d + geom_hex()
```

---

## EDA-SCATTER-SkinThickness vs Insulin
```{r}
scatter2=xyplot(diabetes$SkinThickness ~ diabetes$Insulin, groups=diabetes$Outcome, data = diabetes,auto.key = list(corner = c(0, .98)), cex = .75)
```

---

```{r}
scatter2 ## insulin vs skinThi
```

---

```{r}
d <- ggplot(diabetes, aes(Insulin,SkinThickness))
```

---

```{r}
d + geom_hex()
```

---

## EDA-SCATTER-BMI vs SkinThickness
```{r}
scatter3=xyplot(diabetes$BMI ~ diabetes$SkinThickness, groups=diabetes$Outcome, data = diabetes,auto.key = list(corner = c(0, .98)), cex = .75)
```

---

```{r}
scatter3 ## BMI vs skinthi
```

---

```{r}
d <- ggplot(diabetes, aes(SkinThickness,BMI))
```

---

```{r}
d + geom_hex()
```

---

## EDA-SCATTER-Insulin vs Glucose
```{r}
scatter4=xyplot(diabetes$Glucose ~ diabetes$Insulin, groups=diabetes$Outcome, data = diabetes,auto.key = list(corner = c(0, .98)), cex = .75)
```

---

```{r}
scatter4 ## gluco vs insu
```

---

```{r}
d <- ggplot(diabetes, aes(Insulin,Glucose))
```

---

```{r}
d + geom_hex()
```

---

## EDA-SCATTER-Age vs Glucose
```{r}
scatter5=xyplot(diabetes$Age ~ diabetes$Glucose, groups=diabetes$Outcome, data = diabetes,auto.key = list(corner = c(0, .98)), cex = .75)
```

---

```{r}
scatter5 ## age vs glucose
```

---

```{r}
d <- ggplot(diabetes, aes(Glucose,Age))
```

---

```{r}
d + geom_hex()
```

---

## EDA-HISTOGRAM-Age
```{r}
hist(diabetes$Age,col = "pink", freq = TRUE)
```

---

## EDA-HISTOGRAM-BMI
```{r}
hist(diabetes$BMI,col = "pink", freq = TRUE)
```

---

## EDA-HISTOGRAM-Glucose
```{r}
hist(diabetes$Glucose,col = "pink", freq = TRUE)
```

---

## EDA-HISTOGRAM-BloodPressure
```{r}
hist(diabetes$BloodPressure,col = "pink", freq = TRUE)
```

---

## EDA-BOXPLOT-Glucose
```{r}
boxplot(Glucose~Outcome, data=diabetes)
```

---

## EDA-BOXPLOT-Age
```{r}
boxplot(Age~Outcome, data=diabetes)
```

---

## EDA-BOXPLOT-BMI
```{r}
boxplot(BMI~Outcome, data=diabetes)
```

---

## EDA-BOXPLOT-BloodPressure
```{r}
boxplot(BloodPressure~Outcome, data=diabetes)
```

---

## EDA-BOX&HISTO- AGE
```{r}
plot7={layout(matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,3))
par(mar=c(2, 2, 0.5, 1))
boxplot(Age~Outcome,data=diabetes,horizontal=TRUE,  outline=TRUE, frame=FALSE, col = "green1")
hist(diabetes$Age, col = "blue", freq = FALSE)
lines(density(diabetes$Age), col = "red", lwd = 2)}
```

---

## EDA-BOX&HISTO-BMI
```{r}
layout(matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,3))
par(mar=c(2, 2, 0.5, 1))
boxplot(BMI~Outcome,data=diabetes,horizontal=TRUE,  outline=TRUE, frame=FALSE, col = "green1")
hist(diabetes$BMI, col = "blue", freq = FALSE)
lines(density(diabetes$BMI), col = "red", lwd = 2)
```

---

## EDA-BOX&HISTO- Bloodpressure
```{r}
layout(matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,3))
par(mar=c(2, 2, 0.5, 1))
boxplot(BloodPressure~Outcome,data=diabetes,horizontal=TRUE,  outline=TRUE, frame=FALSE, col = "green1")
hist(diabetes$BloodPressure, col = "blue", freq = FALSE)
lines(density(diabetes$BloodPressure), col = "red", lwd = 2)
```

---

## EDA-BOX&HISTO- Glucose
```{r}
plot7={layout(matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,3))
par(mar=c(2, 2, 0.5, 1))
boxplot(Glucose~Outcome,data=diabetes,horizontal=TRUE,  outline=TRUE, frame=FALSE, col = "green1")
hist(diabetes$Glucose, col = "blue", freq = FALSE)
lines(density(diabetes$Glucose), col = "red", lwd = 2)}
```

---

## EDA-BAR PLOT- Age
```{r}
bar_plt <- ggplot(diabetes, aes(x =Age, fill=Outcome))
bar_plt <- bar_plt + geom_bar()
```

---

```{r}
bar_plt
```

---

## EDA-BAR PLOT- Glucose
```{r}
bar_plt <- ggplot(diabetes, aes(x =Glucose, fill=Outcome))
bar_plt <- bar_plt + geom_bar()
```

---

```{r}
bar_plt
```

---

## EDA-BAR PLOT- BloodPressure
```{r}
bar_plt <- ggplot(diabetes, aes(x =BloodPressure, fill=Outcome))
bar_plt <- bar_plt + geom_bar()
```

---

```{r}
bar_plt
```

---

## EDA-BAR PLOT- BMI
```{r}
bar_plt <- ggplot(diabetes, aes(x =BMI, fill=Outcome))
bar_plt <- bar_plt + geom_bar()
```

---

```{r}
bar_plt
```

---
  
## Data loading for Model prediction
```{r}
rawdata=read.csv("C:/Users/Madan/Documents/diabetes.csv",header=TRUE, stringsAsFactors = TRUE)
rawdata$Outcome <- as.factor(rawdata$Outcome)
set.seed(19910104)
Train <- createDataPartition(rawdata$Outcome, p=0.7, list=FALSE)
nrow(rawdata)
training <- rawdata[ Train, ]
nrow(training)
testing <- rawdata[ -Train, ]
nrow(testing)
```

---
  
## Model 1-GRADIENT BOOSTING-all atributes
```{r}
mod_gbm <- train(Outcome ~ . , data=training, method="gbm")
```

---
  
## Model 1-GRADIENT BOOSTING-all atributes-summary
```{r}
summary(mod_gbm)
```

---
  
## Model 1-GRADIENT BOOSTING-all atributes-varimp
```{r}
varImp(mod_gbm)
```

---
  
## Model 1-GRADIENT BOOSTING-all atributes-Prediction & Accuracy
```{r}
pred_gbm <- predict(mod_gbm,newdata=testing)
confusionMatrix(data=pred_gbm, testing$Outcome)
```

---
  
## Model 1-GRADIENT BOOSTING-selected atributes-training
```{r}
mod_gbm <- train(Outcome ~ Glucose+BMI+Age+Pregnancies+DBF , data=training, method="gbm")
```

---
  
## Model 1-GRADIENT BOOSTING-selected attributes-Prediction & Accuracy
```{r}
pred_gbm <- predict(mod_gbm,newdata=testing)
confusionMatrix(data=pred_gbm, testing$Outcome)
```

---
  
## Model 2-DECISION TREE-training,testing&prediction
```{r}
mod_rpart <- train(Outcome ~ Glucose+BMI+Age+Pregnancies , data=training, method="rpart")
pred_rpart <- predict(mod_rpart,newdata=testing)
confusionMatrix(data=pred_rpart, testing$Outcome)
```

---
  
## Model 3-GENERALIZED LENIAR MODEL-training,testing&prediction
```{r}
mod_rpart <- train(Outcome ~ Glucose+BMI+Age+Pregnancies+DBF , data=training, method="glm")
pred_rpart <- predict(mod_rpart,newdata=testing)
confusionMatrix(data=pred_rpart, testing$Outcome)
```

---
  
## Model 4-NAIVE BAYES-training
```{r}
mod_nb <- train(Outcome ~ Glucose+BMI+Age+Pregnancies, data=training, method="nb")
```

---
  
## Model 4-NAIVE BAYES-prediction
```{r}
pred_nb <- predict(mod_nb, testing, type="raw")
```

---
  
## Model 4-NAIVE BAYES-accuracy
```{r}
confusionMatrix(data=pred_nb, testing$Outcome)
```

---
  
## CONCLUSION
  
1. Model selected according to accuracy
2. Naive bayes has a highest accuracy of 81.3%

---
  
## WEB REFERENCES
  
  1. http://slidify.org/
  2. https://rpubs.com/
  3. https://www.kaggle.com/
  4. https://github.com/
  
---
  
##                   THANK YOU